---
title:  "Project **Brazilian Reefs in the Anthropocene**, Reef Synthesis Working Group  (ReefSYN)"
author: "Dra. Mariana Bender (PI, UFSM), Dr. Sérgio Floeter (Co-PI, UFSC), Dr. André Luís Luza (PostDoc, UFSM), Dr. Linda Eggertsen (PostDoc, UFSM)"
date: "2022/05/05"
output: pdf_document
---



<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
<!-- badges: end -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Who are we?

We are a team of researchers from several countries and institutions, interested in several dimensions of marine biodiversity. We are specially engaged in achieving a common objective: *To provide a synthesis about patterns and drivers of reef diversity and supplying of ecological services in Brazil*. We will synthesize information to be useful to researchers and stakeholders. Our project is funded by the Brazilian CNPq through the Brazilian Institute of Synthesis (Sinbiose, see <http://www.sinbiose.cnpq.br/web/sinbiose/home>)

If you want to know our team and research we do please see <https://reefsyn.weebly.com/home-us.html>.
If you want to know about data please assess <>


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F,fig.align="center", message=F,warning=F}


## R code to load data sets and plot points
## call packages

source("R/packages.R")
source("R/functions.R")

# open researcher data
researcher_data <- read.xlsx(here("Data","researcher_data","PLANILHA_CONSOLIDADA.xlsx"),sheet = 1, colNames = TRUE,detectDates=F)
# formatting data
researcher_data$objetivos <- as.character (researcher_data$objetivos)
researcher_data$titulo_portugues <- as.character (researcher_data$titulo_portugues)
researcher_data$titulo_ingles <- as.character (researcher_data$titulo_ingles)

## extrair as colunas que desejo analaisar
institution <- strsplit(researcher_data$instituicao,",")
institution <- unlist(strsplit(unlist(institution), "/"))
institution <- gsub (" ","",institution)
# df for word-cloud
institution <- data.frame (inst=unique(institution) [order (unique(institution))],
                           freq=table(institution) [order (names(table(institution)))])

### titulo em ingles
titulo_ingles <- as.character (researcher_data$titulo_ingles [which(researcher_data$titulo_ingles != "Indefinido")])
## gerar df para o wcloud
freq_texto_wc <-  function_freq_texto_wc (titulo_ingles)

par(mfrow=c(1,2),mar=rep(0, 4))
## inst
wordcloud(words = institution$inst, freq = institution$freq.Freq,
          min.freq = 1,scale=c(4, .2),
          max.words=50, 
          random.order=FALSE, rot.per=0.15, 
          colors=brewer.pal(8, "Dark2"))

# tit english
wordcloud(words = freq_texto_wc$word, 
          freq = freq_texto_wc$freq.Freq,
          min.freq = 2,scale=c(4, .2),
          max.words=max(freq_texto_wc$freq.Freq)+3, 
          random.order=FALSE, rot.per=0.15, 
          colors=brewer.pal(8, "Dark2"))



```


## Where are we? Which data are we using?

We are analyzing data from 221 unique localities distributed along the Brazilian coast and oceanic islands. In these localities, we gathered published and yet unpublished data of fish and benthic communities --generated by geographically replicated, large-scale research programs conducted over the last decades in Brazilian coast and oceanic islands (SISBOTA-MAR, PELD-ILOC). We are also using data collected beyond the Brazilian coast (e.g., Longo et al. 2019, GEB), and a species-level trait dataset that includes Atlantic and Eastern Pacific fishes (Quimbayo et al., unpublished data).

```{r, echo = F, fig.height=5,fig.width=5,message=F, warning=F}
source("R/packages.R")
source("R/functions.R")


# load folders with data

folders <- list.dirs(here ("DwC_output"))

# load datasets


# benthos

ds1<- read.table (here("DwC_output","AAued_spatialData", "event_core.txt"),sep=",",h=T)
ds2<- read.table (here("DwC_output","GLongo_NRoss_spatialData", "event_core_benthos.txt"),sep=",",h=T)
ds3<- read.table (here("DwC_output","PELD_iloc_benthos", "event_core.txt"),sep=",",h=T)
colnames(ds3)[which(colnames(ds3) == "island")] <- "locationID"
ds3$higherGeographyID <- "BrazilianIslands"
ds4<- read.table (here("DwC_output","RFrancini_spatialData", "event_core.txt"),sep=",",h=T)
ds5<- read.table (here("DwC_output","RFrancini_timeSeries_abrolhos","event_core_benthos.txt"),sep=",",h=T)

# fish

ds6<- read.table (here("DwC_output","GLongo_NRoss_spatialData", "event_core_fish.txt"),sep=",",h=T)
ds7<- read.table (here("DwC_output","GLongo_spatialData", "event_core.txt"),sep=",",h=T)
ds8<- read.table (here("DwC_output","PELD_iloc_fish", "event_core.txt"),sep=",",h=T)
colnames(ds8)[which(colnames(ds8) == "island")] <- "locationID"
ds8 <- ds8 [which(ds8$locationID != "Ascension"),]
ds8$higherGeographyID <- "BrazilianIslands"
ds9<- read.table (here("DwC_output","RFrancini_timeSeries_abrolhos","event_core_fish.txt"),sep=",",h=T)
ds10<- read.table (here("DwC_output","RJ_time_series", "event_core.txt"),sep=",",h=T)
ds11<- read.table (here("DwC_output","RMorais_spatialData", "event_core.txt"),sep=",",h=T)
ds12<- read.table (here("DwC_output","SC_time_series", "event_core.txt"),sep=",",h=T)

#
list_benthos <- list(ds1,ds2,ds3,ds4,ds5)

# agggregate and get coordinates for each site
require(dplyr)
coords <- lapply (list_benthos, function (i) 
  
  
  group_by(i,higherGeographyID, locationID, locality) %>% 
  
                    summarise(decimalLatitude = mean( decimalLatitude),
                            decimalLongitude = mean( decimalLongitude)) %>% 
    
                  select (higherGeographyID,locationID, locality,
                          decimalLatitude,decimalLongitude)
)
# melt
coords <- do.call(rbind,coords)


# mapa mundi
world <- ne_countries(scale = "medium", returnclass = "sf")

# cortar o mapa para ver a america do Sul e parte da central
wm <- ggplot() + 
  geom_sf (data=world, size = 0.1, 
           fill= "gray90",colour="gray90") +
  coord_sf (xlim = c(-50, -25),  ylim = c(-30, 4), expand = FALSE) +
  theme_bw() + xlab ("Longitude")  + ylab ("Latitude") +
  ggtitle ("Data of benthic and fish commnities") + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "lightcyan",#darkslategray1
                                        colour = "lightcyan"),
        
        axis.text.x = element_text(size=6),
        axis.ticks.x=element_line(size=1),
        axis.text.y = element_text(size=6),
        axis.ticks.y=element_line(size=1),
        axis.title.x = element_text(size=8),
        axis.title.y = element_text(size=8),
        title = element_text(size=9)) 

# map
wm_data <- wm +  geom_point(data = coords, aes(x = decimalLongitude, y = decimalLatitude), size =4, 
        shape = 19, col = "green4",alpha=0.3)

wm_data







### map of points with data
```

# Brief description of each data set

## Spatial data

### Fish communities from the Brazilian province
This dataset, used by Morais et al. 2017 (Journal of Fish Biology), includes fish counts in 4,570 transects distributed over 137 sites in 20 different localities from 0 to 27°S latitude degrees (including oceanic islands). The dataset includes information about fish diet, feeding mode, size, and biomass. Sampling descriptors include Observer ID, site depth, and date. The geographical information is at the site (not transect) level. Data and metadata are available [here](https://drive.google.com/drive/u/0/folders/1Wn4n66sVPoj3YBT1kGFUE2FuMyPq4Iic). These data were generated by the team of [SISBIOTA Mar project](http://www.sisbiota.ufsc.br/). All documentation describing the edits in this dataset (for use of the ReefSYN) can be found [here] ().


### Benthic communities from the Brazilian province
This dataset, used by Aued et al. 2018 (PlosOne), includes the cover of benthic organisms into 3,855 photoquadrats deployed in 40 sites in 15 different localities from 0 to 27°S latitude degrees. Benthic organisms were identified at several taxonomic levels (morphotype, species, order). Sampling descriptors include photoquadrat ID, site depth, date and, for some samples, observer ID. The geographical information is at the site (not transect) level. Data and metadata are available [here](https://drive.google.com/drive/u/0/folders/1sgILttlP6HQ1KOHV5yuyQouLe6Pscjlj). These data were generated by the team of [SISBIOTA Mar project](http://www.sisbiota.ufsc.br/).


### Benthic communities from the Brazilian province (Ronaldo Francini-Filho)
This dataset was compiled by Anaide Aued and Erika Santana, and includes the cover of benthic organismos into the photoquadrats sampled by Aued et al. (2018), as well as by the team of Francini-Filho. These data include a subset of taxa included in Aued et al. (2018). Benthic organisms were identified at several taxonomic levels (morphotype, species, order). Sampling descriptors include site depth, month and year. Also include site covariates, such as distance offshore, clorophyll, nitrate, salinity. The geographical information is at the site (not transect) level. Data and metadata are available [here](https://drive.google.com/drive/u/0/folders/1o0S5-u1GAPSLwg8P5zzl11xzx4X5pxIU).

### Trophic interactions along the Western Atlantic
This dataset, used by Longo et al. 2019 (Global Ecology and Biogeograhy), includes records of the interactions between fishes and benthic organisms. These data were obtained with 1,133 unique videoplots deployed in 70 sites in 17 different localities from 34°N to 27°S. Sampling descriptors include recording time, date, depth, and observed ID. No geographical information besides site ID is available. Data and metadata are available [here](https://drive.google.com/drive/u/0/folders/126xLZ6Z3xKO_pIT1IfGNj268Qy7E_Lo7). 

### Fish and benthic communities from Rio Grande do Norte
Data collected by Guilherme Longo and Natália Roos in Rio Grande do Norte. Data of fishes and benthos were collected in the same spatial unit: within a transect of 20m2 used to register fishes, they deployed 10 photoquadrats to register benthic organisms. These data were used by N Roos [Roos et al. 2019](https://www.int-res.com/prepress/m13005.html) and are available [here](https://drive.google.com/drive/u/0/folders/1WG68D_QAPKIZbZhYZmFOenOdUjejnjXn).

## Temporal data
### Benthic communities from oceanic islands
Two datasets of benthic communities registered in the Fernando de Noronha Archipelago and Rocas Atoll from 2013 to 2018. These data were shared by Sérgio Floeter are are available [here](https://drive.google.com/drive/u/0/folders/1DXrcYH3yCbknkEuqpGG59m89LbDJpzVJ). Data from Trindade Island and Martiz Vaz Archipelago, Abrolhos, and Saint Peter and Saint Paul's Archipelago will be available soon. These data were generated by the team of [PELD ILOC project](http://peldiloc.sites.ufsc.br/).

## Trait data
### Atlantic and Eastern Pacific fishes
Unpublished data collected by Juan Pablo Quimbayo and collaborators. These data include functional traits of 2,153 fish species. Traits describe activity, size, diet, distribution, and fishing price. Data and metadata can be found [here](https://drive.google.com/drive/u/0/folders/1aG6zlU5uoeLEPwwdnRKoTAqvldBmyJsp).

### Benthos from the Brazilian province
Unpublished data collected by Anaide Aued and collaborators. These data include functional traits of more than 100 taxa of benthic organisms. Traits describe mobility, size, growth form, trophic level, and accretion capacity. Data and metadata can be found [here](https://drive.google.com/drive/u/0/folders/1H54FZnBRwdF-ZdDkJ7hjB4ZnyBULspV8).

### Corals of the Brazilian province 
Unpublished data collected J Bleuel and collaborators. These data include functional traits of 24 coral species. Traits describe reproduction mode and period, growth rate and form, and tolerance to depth, light, and temperature. These data can be found [here](https://drive.google.com/drive/u/0/folders/1vQp0xz_AMqepVbg1sbrFk1lvsjo_ecYL).    

## Spatial polygons
We are using the most updated, yet unpublished spatial data of marine protected and priority areas (year 2018), from Brazilian Institute for Biodiversity Conservation ('ICMBIO'). These data contain six shapefiles with spatial data about conservation priority areas, artisanal fishing, and implementation cost, urgency, and opportunity. These data were produced by the team of AP Prates, were shared by Carlos Eduardo Ferreira, and can be found [here](https://drive.google.com/drive/u/0/folders/1Xra9owUYi0dK_T4l8WEn-aH7QnTk6ZtH).


```{r,fig.height=7, fig.width=7, echo= F,message=F,warning=F, highlight=F, results= "hide"}



```

## Brazilian Protected Areas

Finally, we have spatial data on Brazilian Protected Areas that can be used in many steps of the project. These spatial data were shared by Rafael Magris (ICMBio) and can be found [here] ().

## How can you access these data?

Published data can be found at their on-line repositories. Unpublished data are only available to ReefSyn members, as defined by our "Policy of data sharing and use" (available [here](https://github.com/Sinbiose-Reefs/reefsyn_site/blob/master/DataPolicy_SINBIOSE.pdf)). We will let these data available to external users at **January 2023**, one year after the end of the ReefSyn project. If you are a member of the ReefSyn team, please read our [co-authoring rules](https://github.com/Sinbiose-Reefs/reefsyn_site/blob/master/Coauthor_SINBIOSE.pdf).
  
### Grant

This project is granted by the Brazilian Biodiversity Synthesis Center - [SINBIOSE](http://www.sinbiose.cnpq.br/web/sinbiose). This center belongs to the National Council for Scientific and Technological Development (*Conselho Nacional de Desenvolvimento Científico e Tecnológico*, CNPq).  
    
    
    
    
**ReefSYN Working Group**  
  
  


